\documentclass[runningheads]{llncs}

\usepackage[T1]{fontenc}
\usepackage{array}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{url}

\newcommand{\twolinecol}[2]{\texttt{#1}\newline\texttt{#2}}
\pgfplotsset{compat=1.18}

\title{KNIME Workflows in Scientific Studies:
Publication Practices and Reproducibility Risks}
\titlerunning{KNIME Workflow Publication Practices}

\author{Vitalii Kaplan\orcidID{0009-0009-8181-2863}}
\authorrunning{V. Kaplan}
\institute{\email{2333275007@alu.istinye.edu.tr}}

\begin{document}
\maketitle

\begin{abstract}
In this paper, we study how visual workflows are reported in scientific
publications and how incomplete workflow preservation can make older studies
harder to reproduce as workflow platforms evolve. We use KNIME as a case
study, combining OpenAlex bibliometrics, a 100-record top-cited article audit
registry, workflow retrieval and current-KNIME opening/execution checks, and
repository-level mining of KNIME node metadata. OpenAlex shows that KNIME is
visible across a broad scientific literature, while the article assessment
shows that influential studies often report KNIME through figures,
descriptions, or tool mentions rather than downloadable workflow artifacts. In
the 100-record article audit registry, twenty-two records report
downloadable or linked KNIME workflow files. Recorded retrieval attempts
obtained workflow artifacts or workflow-directory evidence for eighteen article records. In the recorded
manual execution checks, five article records had at least one workflow execute
successfully.
Repository mining gives the source-evolution context: deprecated ordinary
nodes increased from 14.83\% of registered ordinary nodes in the 2018 source
baseline to 33.33\% in the June 28, 2026 local source snapshot. Together,
these results show that weak workflow preservation and evolving node metadata
can make long-term reproduction of KNIME-based studies harder over time.

\keywords{Workflow reproducibility \and KNIME \and Deprecated nodes \and
Software evolution}
\end{abstract}

\section{Introduction}

KNIME is a widely used open-source platform for constructing scientific
data-analysis workflows, with more than a decade of use across published
studies. Its visual workflows represent analyses as connected nodes with
stored settings, data ports, and extension dependencies, which can make
computational work easier to inspect, reuse, and repeat. This representation
provides more operational detail than a narrative method description, but it
does not remove the need to preserve the executable workflow artifact and its
context. Reproduction requires access to the workflow file, KNIME version,
extensions, data, and execution environment. Figures and textual summaries do
not provide sufficient substitutes for these materials.

This issue matters for visual workflow platforms in general, because their
reproducibility value depends on preserving executable workflow artifacts and
the software context needed to interpret them. We study KNIME because it is an
open-source visual workflow platform with a long public development history and
substantial visibility in scientific literature. Our OpenAlex query returned
963 article-type records whose titles or abstracts match \texttt{KNIME}. The
matching records span platform papers, extension papers, tool-comparison
papers, and domain studies that use KNIME as an analysis interface. KNIME is
therefore a suitable case for studying whether visual workflows are preserved
as executable artifacts or only reported through descriptions and figures.

Our top-cited article collection covers the 100 most-cited KNIME-matching
records, with not-assessed placeholders retained where local full text is
unavailable. The structured audit reported in
Table~\ref{tab:top-cited-assessment} shows that publication practice is the
weak link. Twenty-two records report downloadable or linked KNIME workflow
files, and workflow artifacts or workflow-directory evidence were obtained for eighteen article records in
the recorded retrieval attempts. Other links were stale, blocked, incomplete,
or did not lead to a confirmed KNIME workflow artifact. In the recorded manual
execution checks, five article records had at least one workflow execute
successfully. For other influential studies, the strongest workflow evidence
is usually weaker than an executable artifact. Several records present
screenshots or figures of connected KNIME nodes, or describe workflow steps in
text, without preserving all executable workflow context.

These reporting practices are risky because KNIME continues to evolve. A
workflow that was executable when a paper was written may later depend on nodes
that are deprecated, hidden, migrated, removed, or supplied by third-party
extensions that are no longer easy to install. This paper therefore connects
publication evidence with platform-evolution evidence. We combine OpenAlex
bibliometrics, top-cited article retrieval and assessment, workflow-artifact
retrieval and current-version execution attempts, and repository-level mining
of KNIME node metadata. The goal is to show why publication practice and
source-code evolution must be considered together when evaluating long-term
reproducibility of KNIME-based studies.

The rest of the paper is organized as follows. Section~2 reviews related work
on computational reproducibility, open-source workflow systems, environment
preservation, and platform evolution. Section~3 presents the methods before
the data because the study depends on how the bibliometric records, article
assessments, and repository snapshots were collected. Section~4 describes the
resulting datasets. Section~5 reports the bibliometric, article-audit,
workflow-retrieval, manual compatibility-test, and repository-mining results.
Section~6 discusses
the implications, limitations, and future extensions of the study.

\section{Literature Review}

Reproducibility is a central requirement for computational research because
published results increasingly depend on software, data transformations, and
execution environments. Provenance research makes this requirement explicit:
visual analytics systems need records of data, analytical processes, and human
interpretation to make later inspection and reuse possible
\cite{Walker2013ProveML}. Studies of executable research artifacts show the
same problem in practice. Samuel and Mietchen found that biomedical Jupyter
notebooks often fail during dependency installation or re-execution even when
the notebooks themselves are available
\cite{SamuelMietchen2024JupyterReproducibility}.

Open-source scientific software helps reproducibility by making algorithms,
interfaces, and execution infrastructure inspectable. Weka, for example, is
presented as an extensible open-source data-mining environment with graphical
interfaces, reusable configurations, and plugin mechanisms
\cite{Hall2009WEKA}. KNIME is similarly used as an open-source visual workflow
platform, and recent GIScience work explicitly proposes KNIME-based workflows
as an approach for advancing replicable and reproducible research
\cite{Fu2025GIScienceKNIME}. However, openness of the platform alone is
insufficient. A workflow-based study also requires the executable workflow
file, input data, platform version, installed extensions, and other software
dependencies. Container research makes the same point at the environment
level: scientific computations are easier to reproduce when complete software
environments can be preserved and moved between systems
\cite{Kurtzer2017Singularity}. More broadly, the testing literature shows that
executable artifacts can remain sensitive to operating systems, dependency
versions, timing, external services, and other environmental factors
\cite{Rasheed2023FlakinessReview}.

Even when an executable workflow is published, long-term reproduction can be
affected by the evolution of the workflow platform itself. KNIME's node extension
schema states that a deprecated node is no longer shown in the node repository,
but is still loaded in existing workflows
\cite{KNIMEWorkbenchNodeSchema2026}. This design supports backward
compatibility, while also marking a change in the ordinary authoring surface.
KNIME workflows also persist node factory class names, and the platform
provides extension points for mapping old factory classes to newer
implementations and for registering workflow-node migration rules
\cite{KNIMECoreClassMapperSchema2026,KNIMEMigrationRuleSchema2026}. These
mechanisms show that continued workflow loading may depend on compatibility
metadata, extension availability, and migration support. We therefore treat
KNIME node deprecation as a measurable risk marker for long-term workflow
reproducibility.

\section{Methods}

All scripts and data used for this study, except the other articles assessed
as study objects, are available in the public replication repository
\cite{KaplanReproducibilityRisksArticle2026}. Relative paths reported in this
paper should be interpreted as paths within that GitHub repository.

\subsection{Bibliometric Evidence}

We collected bibliometric evidence from the OpenAlex Works API
\cite{Priem2022OpenAlex} on June 29, 2026. The query used OpenAlex
title-and-abstract search for the word \texttt{KNIME} and restricted the
result set to OpenAlex records whose type is \texttt{article}. The same query
can be reproduced in the
OpenAlex web interface by selecting ``Search title and abstract'', entering
\texttt{KNIME}, and filtering the result type to \texttt{article}. The
corresponding web-interface request can be represented as:

\begin{quote}
\path{https://openalex.org/works}\\
\path{page=1}\\
\path{sort=relevance_score:desc}\\
\path{search.title_and_abstract=KNIME}\\
\path{filter=type:article}
\end{quote}

The first API request, which returns the first 200 records, is reproducible as:

\begin{quote}
\path{https://api.openalex.org/works}\\
\texttt{?search.title\_and\_abstract=KNIME}\\
\texttt{\&filter=type:article}\\
\texttt{\&per-page=200}\\
\texttt{\&cursor=*}
\end{quote}

OpenAlex uses cursor pagination for large result sets. After the first request,
the collector reads \texttt{meta.next\_cursor} from each response and submits
the same request with \texttt{cursor=<next-cursor>} until no further results
are returned. The value \texttt{per-page=200} is the maximum page size used
here.

The collection script is
\path{scripts/collect_openalex_knime_works.py}.
It stores raw and processed collection artifacts as follows:
\begin{itemize}
\item raw OpenAlex response pages:
\path{data/original/openalex/pages/}
\item complete record stream:
\path{data/original/openalex/works.jsonl}
\item compact CSV view:
\path{data/original/openalex/works.csv}
\item endpoint, query parameters, collection time, result count, and page
manifest:
\path{data/original/openalex/metadata.json}
\end{itemize}
The current run collected 963 records across 5 OpenAlex response pages.
Derived bibliometric tables are generated from the JSONL file by
\path{scripts/build_openalex_knime_bibliometrics.py}.

\subsection{Top-Cited Article Assessment}

For the article-level audit we developed an open-source application composed
of scripts, JSON configuration files, and prompt configuration files. The
process is specified in \path{scripts/audit_script_chain.json}, which acts as
the workflow index for this part of the study. Each step consumes the recorded
outputs of earlier steps and writes an explicit intermediate or final artifact.
The current chain contains fifteen scripted steps and one manual checkpoint.
Thirteen scripted steps are rule-based. Two scripted steps use an LLM.
We use the LLM only where rule-based extraction is insufficient. The first LLM
step classifies workflow relevance from downloaded reference-page content. The
second LLM step fills supplementary article-level audit flags from the
processed text of the articles. The LLM calls use temperature 0 to reduce
variation. Repeated execution may still depend on the model and service version
used.

The pipeline is mostly automatic, but it is deliberately semi-automatic at two boundary
points: obtaining article PDFs at the beginning of the process and obtaining
workflow artifacts near the end, before formal result preparation.
Both tasks may involve institutional access, protected publisher pages,
supplement portals, captchas, and changing web interfaces. The process can
therefore be improved by human assistance at the beginning and end, while the
middle steps are executed automatically from the recorded intermediate files.

The implementation was run with Python 3.14.5 and uses a separate
\path{.venv-playwright} virtual environment for browser-dependent steps.
Article PDF conversion uses GROBID \texttt{0.9.0-crf} through the Docker image
\texttt{grobid/grobid:0.9.0-crf} \cite{GROBIDSoftware2026}. Browser-assisted
fetching uses Playwright for Python 1.61.0 with Chromium/Chrome
\cite{PlaywrightPython2026}. The two LLM-supported steps call the OpenAI Chat
Completions API directly, using the configured model
\texttt{gpt-4.1-mini} and temperature 0 \cite{OpenAIAPI2026}. OpenAlex is the
bibliographic source for article discovery and ranking
\cite{Priem2022OpenAlex}. The exact commands, script order, prompt files, and
intermediate outputs are recorded in \path{scripts/audit_script_chain.json}.

For reproducibility of the LLM-supported steps, API configuration is separated
from prompts and data. The scripts read the API key from \path{.env} or the
process environment as \texttt{OPENAI\_API\_KEY}; this local credentials file
is not part of the research data. The model can be overridden with
\texttt{OPENAI\_MODEL}, but the recorded run used the default
\texttt{gpt-4.1-mini}. Both LLM scripts use the Chat Completions endpoint
\path{https://api.openai.com/v1/chat/completions}, temperature \texttt{0.0},
and at most three retries for transient server-side failures. The prompts are
stored as JSON configuration files:
\path{data/processed/audit/article_reference_classification_prompt.json} for
reference-page classification and
\path{data/processed/audit/article_supplementary_flag_prompt.json} for
article-level supplementary flags. The derived JSON outputs record the script
name, input file, prompt configuration, model, temperature, selection scope,
summary counts, and per-article classifications. This makes the exact prompt
and local evidence auditable, while recognizing that future reruns may differ
if the hosted model implementation changes.

Step 1 collects KNIME-matching article records from the OpenAlex works API with
\path{scripts/collect_openalex_knime_works.py}. The query searches for KNIME in
article titles and abstracts and stores the original API response pages, a
complete JSONL stream, a compact CSV view, and query metadata. This step
preserves the bibliographic source records before any article-level audit
decisions are made.

Step 2 builds the bibliometric summaries and the citation-ranked seed table
with \path{scripts/build_openalex_knime_bibliometrics.py}. The script reads the
OpenAlex JSONL export and derives processed CSV and JSON files used to select
the top-cited audit candidates. The citation-ranked seed table is then used as
the article list for downstream download, parsing, URL collection, and audit
report construction.

Step 3 attempts to download article PDFs by citation-rank range with
\path{scripts/download_openalex_rank_range_articles.py}. This step is
semi-automatic because OpenAlex links and open-access URLs do not always expose
the full paper, and some papers require lawful institutional access. The script
records download attempts and stores retrieved PDFs under
\path{data/original/articles/}; failed or protected records can be completed by
manual download without changing the later audit logic.

Step 4 converts the local PDFs into semantic article HTML with
\path{scripts/extract_article_grobid_html.py}. The current parser uses a local
GROBID service to produce TEI XML and then renders an HTML reading copy with
article metadata, abstract, body sections, paragraphs, and bibliography
structure. This replaced earlier plain-text extraction because the downstream
steps need stable article identity, paragraphs, and reference context.

Step 5 collects article metadata and article-related URLs with
\path{scripts/collect_article_urls.py}. The script joins the OpenAlex seed
table, the canonical PDF registry, the GROBID TEI files, and the generated HTML
files. At this stage it does not classify URLs or infer workflow availability;
it only records metadata and normalized URLs that can be traced to the article
content.

Step 6 fetches the referenced URL pages with
\path{scripts/fetch_article_url_pages.py}. For each unique URL, the script
records HTTP status, redirect chain, final URL, response headers, and a stored
copy of the response body where available. This creates local evidence for
later classification and avoids asking the language model to rely on live web
state.

Step 7 attaches the URL-fetch metadata back to the article records with
\path{scripts/attach_url_fetch_metadata.py}. This rule-based step augments
each URL entry with fetch status, HTTP code, final URL, redirect count, and the
path to any stored page. It is metadata attachment only; it does not interpret
whether a reference is a workflow, software page, dataset, article landing
page, or documentation.

Step 8 performs a second-stage fetch for reference pages that ordinary HTTP
requests cannot access, using \path{scripts/browser_fetch_403_urls.py}. This
step uses a browser when useful, stores rendered pages, and records challenge
or blocked states explicitly. The process
does not solve captchas or bypass access controls; protected pages are treated
as evidence of access friction rather than forced downloads.

Step 9 classifies reference pages for workflow obtainability with
\path{scripts/classify_article_references_with_llm.py}. The language model is
given URL metadata and locally cached reference-page excerpts, not the whole
article as its primary evidence. The purpose is to decide whether a reference
can help a reader obtain a KNIME workflow artifact, whether it only points to
software, documentation, data, or a general page, or whether manual checking is
needed.

Step 10 classifies supplementary article-level flags with
\path{scripts/classify_article_supplementary_flags_with_llm.py}. This LLM step
uses the article HTML, metadata, and reference classifications to fill bounded
boolean fields comparable to a manual audit checklist: KNIME use, KNIME version
reporting, workflow screenshots or figures, textual workflow or node
descriptions, input-data availability, code or scripts, and extension or
plugin information. These flags provide article-level context around the
reference-page workflow evidence.

Step 11 builds a compact article audit report with
\path{scripts/build_article_audit_report.py}. This rule-based join combines
article metadata, the supplementary flag fields, and the workflow-relevant
reference classifications into
\path{data/processed/audit/article_audit_report.json}. The report is the
central article-level audit artifact used by later workflow-download and table
generation steps.

Step 12 attempts automatic workflow artifact downloads from the workflow
references in the audit report with
\path{scripts/download_workflow_artifacts.py}. The script uses rule-based
HTTP retrieval and, where appropriate, browser-assisted discovery to look for
KNIME workflow archives, \texttt{workflow.knime} files, or archives containing
workflow files. It writes attempt status, local directories, downloaded files,
and discovered workflow files to the workflow-reference inventory.

Step 13 is the manual workflow-download checkpoint. The auditor inspects
\path{data/processed/audit/knime_downloadable_workflow_references.json} and
tries to download missing workflow artifacts where access is allowed. The
inventory JSON is not edited manually at this stage; the manual action is to
place obtained files in the article-specific directory under
\path{data/original/workflows/}. This preserves a clear boundary between human
retrieval and scripted evidence updating.

Step 14 updates the article report from retrieved workflow evidence with
\path{scripts/update_article_report_from_workflow_inventory.py}. The script scans
the workflow inventory and the actual files under \path{data/original/workflows/}
and updates the article audit report when a downloadable KNIME workflow has
been obtained. This makes the report reflect the local evidence rather than
only the earlier URL classification.

Step 15 generates the article-audit summary table source with
\path{scripts/build_article_audit_tables.py}. The script reads the compact
article audit report and writes the CSV files used for the publication-level
summary table. The table data are therefore derived from the final structured
report rather than maintained by hand.

Step 16 generates the KNIME-use workflow-reporting table source with
\path{scripts/build_knime_use_workflow_reporting_table.py}. This step combines
the article-level audit report with the workflow-reference inventory, so that
publication-level workflow-reporting signals and actual obtained workflow
evidence are kept separate but can be summarized consistently.

The resulting process is reproducible at the level of inputs, scripts,
configuration files, cached pages, LLM prompts, and derived JSON reports, while
still acknowledging that publication and workflow retrieval are not fully
automatable. Manual actions are restricted to acquiring protected article PDFs
and workflow artifacts where access is lawful; the subsequent interpretation,
report updating, and table construction are scripted from recorded evidence.

\subsection{Repository-Level Analysis}

We study KNIME node deprecation through repository mining of the public
\texttt{knime-oss} GitHub organization. KNIME Analytics Platform is a
workflow-based data-analysis system in which analyses are assembled from
connected processing nodes
\cite{Berthold2009KNIME,KNIMEUserGuide2026}. KNIME extensions are distributed
through update sites and extend the platform with additional functionality
\cite{KNIMEExtensions2026}. The KNIME source repositories were cloned locally,
and each source snapshot was reconstructed by checking out every non-hidden
top-level Git repository to the latest commit at or before a selected target
date. This checkout step is implemented by
\path{scripts/checkout_knime_oss_by_date.sh}.
The script records a manifest for each snapshot, including repository name,
checkout status, target date, selected commit, selected commit date, and
previous head.

The current repository-mining corpus contains 91 immediate Git repositories in
the local clone. The date-based mining process includes 90 non-hidden immediate
repositories. The hidden \texttt{.github} repository is excluded because it
contains organization metadata rather than KNIME node implementation code. The
exact repository list is stored in
\path{data/processed/knime_snapshots/}
\path{knime_oss_repositories.csv}.

For each date-based source snapshot, we parse KNIME metadata files rather than
grepping source text. The main deprecation marker is
\texttt{deprecated="true"} on ordinary node registrations in
\texttt{plugin.xml} under the Eclipse extension point
\begin{quote}
\texttt{org.knime.workbench.repository.nodes}
\end{quote}
The KNIME schema states that deprecated nodes are no longer shown in the node
repository but are still loaded in existing workflows
\cite{KNIMEWorkbenchNodeSchema2026}. Dynamic node sets use a separate
extension point:
\begin{quote}
\texttt{org.knime.workbench.repository.nodesets}
\end{quote}
We count these registrations separately. Node-description XML files ending in
\texttt{NodeFactory.xml} and rooted at \texttt{knimeNode} are parsed as a
secondary deprecation source.
Hidden nodes, marked with \texttt{hidden="true"}, are counted separately from
deprecated nodes.

The extraction also records compatibility-related metadata from two extension
points:
\begin{quote}
\texttt{org.knime.core.NodeFactoryClassMapper}
\end{quote}
\begin{quote}
\texttt{org.knime.workflow.migration.NodeMigrationRule}
\end{quote}
Together, these extension points document mechanisms for mapping persisted
node-factory class names to newer implementations and for registering
workflow-node migration rules
\cite{KNIMECoreClassMapperSchema2026,KNIMEMigrationRuleSchema2026}. These
records are not counted as deprecation markers, but they are relevant for later
analysis of whether deprecated or replaced nodes have explicit migration
support.

The parser is implemented in
\path{scripts/collect_knime_node_snapshot.py}.
It uses Python's standard XML parser and skips \texttt{.git},
\texttt{target}, \texttt{bin}, and \texttt{.metadata} directories. Boolean XML
attributes are treated as true only when their value is case-insensitively
equal to \texttt{true}.

For each snapshot, the extractor writes per-record tables for node
registrations, node descriptions, factory-class mappers, migration rules, and a
local summary. The per-snapshot outputs are stored under
\path{data/original/knime_snapshots/<snapshot-date>/}
and include:
\begin{itemize}
\item \path{logs/checkout_<snapshot-date>.csv}
\item \path{plugin_nodes.csv}
\item \path{node_descriptions.csv}
\item \path{factory_class_mappers.csv}
\item \path{migration_rules.csv}
\item \path{summary.csv}
\end{itemize}
The cross-snapshot table is stored as:
\begin{quote}
\path{data/processed/knime_snapshots/}

\path{knime_node_snapshot_summary.csv}
\end{quote}
It is generated from these per-snapshot outputs by
\path{scripts/build_knime_node_snapshot_summary.py}.

The summary builder aggregates ordinary node registrations, dynamic nodesets,
unique factory classes, deprecated nodes, hidden nodes, node-description files,
deprecated node descriptions, factory-class mappers, and migration rules. It
also counts processed and skipped repositories from each checkout manifest.
Transition columns compare adjacent chronological snapshots. Node identity is
based on \texttt{factory\_class} where available. Otherwise, the fallback key
is \texttt{plugin\_xml:element:category\_path}. These transition counts are
therefore metadata-level approximations and are strongest for nodes with stable
factory classes.

All project Python scripts are intended to be run with Python 3.11 or newer.
The current scripts use only the Python standard library. No third-party
Python packages are required for the OpenAlex and KNIME repository-mining
pipelines.

\section{Data}

\subsection{Bibliometric Evidence}

The OpenAlex bibliometric data are organized into original API responses and
processed analysis tables. The original data are stored under
\path{data/original/openalex/}.
They include raw response pages, a complete JSONL export, a compact CSV export,
and query metadata. These files preserve the article-level records used to
measure the size, time trend, venues, fields, and accessibility of the
KNIME-matching literature.

Derived bibliometric tables are stored under
\path{data/processed/openalex/}.
Their roles in the analysis are:

\begin{itemize}
\item \path{openalex/openalex_knime_articles.csv}: compact article-level
metadata used as the main processed bibliography table.
\item \path{openalex/openalex_knime_articles_by_year.csv}: annual publication
counts used to describe the time trend of KNIME-matching literature.
\item \path{openalex/openalex_knime_top_venues.csv}: source-level aggregation
used to identify journals, proceedings, repositories, and platforms where the
matching records appear.
\item \path{openalex/openalex_knime_top_domains.csv}: OpenAlex domain
aggregation used to describe the broad disciplinary distribution.
\item \path{openalex/openalex_knime_top_fields.csv}: OpenAlex field
aggregation used to describe the finer disciplinary distribution.
\item \path{openalex/openalex_knime_most_cited.csv}: citation-ranked subset
used to select influential records for later manual assessment.
\item \path{openalex/openalex_knime_likely_workflow_platform.csv}: heuristic
subset used to prioritize papers likely to use KNIME as a workflow platform.
\end{itemize}

The processed OpenAlex directory also contains additional summary files:
\begin{itemize}
\item \path{openalex/openalex_knime_open_access_availability.json}
\item \path{openalex/openalex_knime_role_classification_summary.json}
\end{itemize}
These files summarize artifact-access signals and role-classification counts.

\subsection{Top-Cited Article Assessment}

The top-cited article dataset records the structured assessment of the 100
most-cited OpenAlex records. The citation-ranked source list is
\path{data/processed/openalex/openalex_knime_most_cited.csv}, and the local
article registry is stored under \path{data/original/articles/}. Local PDFs,
where available, are preserved in that directory, while download-attempt logs
are stored under \path{data/processed/audit/logs/}.

Article text and references are represented by GROBID outputs. Semantic HTML
files are stored under \path{data/processed/articles/}, and TEI XML files are
stored under \path{data/processed/articles/grobid_tei/}. These outputs provide
article text, section structure, bibliography context, and machine-readable
links for the later URL and LLM steps.

The URL extraction step writes
\path{data/processed/audit/article_url_collection.json}. For each article, this
file stores rank, canonical identifiers, PDF path, TEI path, HTML path,
OpenAlex metadata, and article-related URLs extracted from the GROBID outputs.
The fetched reference-page cache is stored under
\path{data/processed/audit/pages/}; browser-rendered fetches for selected
blocked pages are stored under \path{data/processed/audit/browser_pages/}.
These caches include HTTP status, redirects, final URLs, headers, stored body
paths, and fetch errors or challenge pages where applicable.

The LLM-derived article-audit data are split into two files.
\path{data/processed/audit/article_reference_llm_classifications.json}
classifies cached article references by workflow obtainability and records the
page evidence used for those classifications.
\path{data/processed/audit/article_supplementary_llm_flags.json} records
article-level flags for KNIME use, version reporting, workflow figures,
workflow text descriptions, data availability, code availability, and
extension information.

The compact report used for the article-level tables is
\path{data/processed/audit/article_audit_report.json}. It merges article
metadata, supplementary flags, and workflow-relevant reference classifications
into one record per article. Workflow-download attempts and local workflow
paths are recorded in
\path{data/processed/audit/knime_downloadable_workflow_references.json}; manual
KNIME-opening screenshots are stored under each workflow directory's
\path{opened/} subdirectory. Table sources are generated under
\path{article/tables/} from these JSON files rather than edited manually.

\subsection{Repository-Level Analysis}

The KNIME repository-mining data are organized as original per-snapshot audit
files and one processed cross-snapshot summary. The processed summary is stored
as:
\begin{quote}
\path{data/processed/knime_snapshots/}

\path{knime_node_snapshot_summary.csv}.
\end{quote}
Per-snapshot audit files are stored under
\path{data/original/knime_snapshots/<snapshot-date>/} and include:

\begin{itemize}
\item \path{logs/checkout_<snapshot-date>.csv}
\item \path{plugin_nodes.csv}
\item \path{node_descriptions.csv}
\item \path{factory_class_mappers.csv}
\item \path{migration_rules.csv}
\item \path{summary.csv}
\end{itemize}

\begingroup
\small
\begin{longtable}{%
  >{\raggedright\arraybackslash}p{0.40\textwidth}
  >{\raggedright\arraybackslash}p{0.18\textwidth}
  >{\raggedright\arraybackslash}p{0.32\textwidth}}
\caption{Columns in the processed KNIME node snapshot summary.}
\label{tab:summary-columns}\\
\toprule
Name & Type & Short description \\
\midrule
\endfirsthead
\toprule
Name & Type & Short description \\
\midrule
\endhead
\bottomrule
\endfoot
\texttt{snapshot\_id} & string &
Stable identifier for the snapshot. \\
\texttt{snapshot\_kind} & categorical &
Snapshot basis, currently date-based. \\
\texttt{snapshot\_date} & date &
Target date used for repository checkout. \\
\texttt{knime\_version} & string &
Version label or source-date context assigned by the summary script. \\
\texttt{source\_basis} & categorical &
Method used to obtain the source snapshot. \\
\texttt{repos\_processed} & integer &
Repositories with a checkout commit in the snapshot manifest. \\
\texttt{repos\_skipped} & integer &
Repositories without a commit at or before the target date. \\
\texttt{plugin\_xml\_files} & integer &
Distinct parsed \texttt{plugin.xml} files contributing node metadata. \\
\texttt{repos\_with\_plugin\_xml} & integer &
Repositories contributing at least one parsed \texttt{plugin.xml}. \\
\texttt{registered\_nodes} & integer &
Ordinary node registrations in the node extension point. \\
\texttt{registered\_nodesets} & integer &
Dynamic nodeset registrations in the nodeset extension point. \\
\texttt{registered\_total} & integer &
Sum of ordinary node and nodeset registrations. \\
\texttt{unique\_factory\_classes} & integer &
Distinct non-empty factory classes among ordinary node registrations. \\
\texttt{deprecated\_nodes} & integer &
Ordinary nodes with \texttt{deprecated="true"}. \\
\texttt{deprecated\_nodesets} & integer &
Nodesets with \texttt{deprecated="true"}. \\
\texttt{deprecated\_total} & integer &
Sum of deprecated ordinary nodes and deprecated nodesets. \\
\twolinecol{unique\_deprecated}{\_factory\_classes} & integer &
Distinct factory classes among deprecated ordinary nodes. \\
\texttt{deprecated\_node\_percent} & percent &
Deprecated ordinary nodes divided by registered ordinary nodes. \\
\texttt{hidden\_nodes} & integer &
Ordinary nodes with \texttt{hidden="true"}. \\
\texttt{hidden\_node\_percent} & percent &
Hidden ordinary nodes divided by registered ordinary nodes. \\
\twolinecol{deprecated\_and}{\_hidden\_nodes} & integer &
Ordinary nodes marked both deprecated and hidden. \\
\texttt{node\_description\_files} & integer &
Parsed \texttt{*NodeFactory.xml} files rooted at \texttt{knimeNode}. \\
\twolinecol{description\_deprecated}{\_files} & integer &
Node-description files with \texttt{deprecated="true"}. \\
\twolinecol{description\_deprecated}{\_percent} & percent &
Deprecated node descriptions divided by parsed node descriptions. \\
\twolinecol{factory\_class\_mapper}{\_count} & integer &
Registered \texttt{NodeFactoryClassMapper} contributions. \\
\texttt{migration\_rule\_count} & integer &
Registered \texttt{NodeMigrationRule} contributions. \\
\twolinecol{nodes\_added}{\_since\_previous} & integer/blank &
Node identity keys present in this snapshot but not the previous one. \\
\twolinecol{nodes\_removed}{\_since\_previous} & integer/blank &
Node identity keys present in the previous snapshot but absent here. \\
\twolinecol{nodes\_newly\_deprecated}{\_since\_previous} & integer/blank &
Common node keys that became deprecated since the previous snapshot. \\
\twolinecol{nodes\_no\_longer\_deprecated}{\_since\_previous} &
integer/blank &
Common node keys no longer marked deprecated. \\
\twolinecol{nodes\_newly\_hidden}{\_since\_previous} & integer/blank &
Common node keys that became hidden since the previous snapshot. \\
\twolinecol{nodes\_no\_longer\_hidden}{\_since\_previous} & integer/blank &
Common node keys no longer marked hidden. \\
\twolinecol{nodes\_category\_changed}{\_since\_previous} & integer/blank &
Common node keys whose observed category-path set changed. \\
\end{longtable}
\endgroup

\section{Results}

\subsection{Bibliometric Evidence}

The OpenAlex title-and-abstract query returned 963 article-type records. This
is the first piece of evidence for the study: KNIME is visible in a sizeable
scientific literature, so preservation of KNIME workflows is not only a niche
platform-maintenance issue. The annual series indicates that KNIME-matching
literature becomes visible in OpenAlex from the mid-2000s and grows into a
sustained publication stream after 2015 (Fig.~\ref{fig:openalex-by-year}). The
highest annual count in the current export is 104 records in 2023. Counts for
2026 are incomplete because the query was collected on June 29, 2026.

\begin{figure}
\centering
\begin{tikzpicture}
\begin{axis}[
  width=0.95\textwidth,
  height=0.34\textwidth,
  ybar,
  bar width=3.5pt,
  xlabel={Publication year},
  ylabel={Article records},
  ymin=0,
  xmin=2005,
  xmax=2027,
  enlarge x limits=0.02,
  xtick={2006,2010,2015,2020,2026},
  ymajorgrids=true,
  grid style={gray!25},
  tick label style={font=\scriptsize},
  label style={font=\small},
]
\addplot table[
  x index=0,
  y index=1,
  col sep=comma,
  ignore chars={\"}
] {../data/processed/openalex/openalex_knime_articles_by_year.csv};
\end{axis}
\end{tikzpicture}
\caption{OpenAlex article records matching \texttt{KNIME} in title or
abstract by publication year.}
\label{fig:openalex-by-year}
\end{figure}

The domain distribution shows that the KNIME-matching literature is not
confined to one disciplinary area (Table~\ref{tab:openalex-domains}). OpenAlex
assigns just over half of the records to Physical Sciences, but the remaining
records are spread across Social Sciences, Life Sciences, and Health Sciences.
This matters for the study because workflow reproducibility risk is not only a
software-engineering concern. If KNIME workflows are used across computational
chemistry, biomedical analysis, education, business analytics, and related
areas, then missing workflow artifacts and changing node implementations can
affect published computational evidence in several research communities.

\begin{table}
\centering
\caption{OpenAlex domains for KNIME title-and-abstract matches.}
\label{tab:openalex-domains}
\begin{tabular}{lrr}
\toprule
Domain & Articles & Share \\
\midrule
Physical Sciences & 505 & 52.44\% \\
Social Sciences & 187 & 19.42\% \\
Life Sciences & 149 & 15.47\% \\
Health Sciences & 122 & 12.67\% \\
\bottomrule
\end{tabular}
\end{table}

The ten largest OpenAlex fields are led by Computer Science (350),
Biochemistry, Genetics and Molecular Biology (115), Medicine (92),
Engineering (71), and Social Sciences (66). These field counts show why a
paper about KNIME reproducibility cannot focus only on the KNIME platform
literature. The central reproducibility question is how scientific studies
that use KNIME as a tool report the executable workflow, data, versions, and
extensions needed for later reproduction.

Open-access metadata indicate 619 open-access records, 625 records with some
full-text availability, and 421 records with a PDF signal. These signals are
important because later workflow audits require access to paper text,
supplements, repositories, and sometimes PDF-only workflow descriptions. The
counts are only availability indicators. They do not by themselves establish
that a workflow artifact, KNIME version, input data, or extension list is
available.

The heuristic role classification separates likely papers about KNIME or KNIME
extensions from papers that likely use KNIME as an analysis tool. The current
rules classify 156 records as about KNIME or a KNIME extension, 314 as using
KNIME as a tool, 28 as about or mentioning KNIME in the title, and 465 as
ambiguous OpenAlex matches. The same heuristic marks 410 records as likely
using KNIME as a workflow platform. These labels are useful for triage: they
identify a substantial candidate set where workflow reporting can matter, but
final case-study claims require manual full-text and artifact assessment.

\subsection{Top-Cited Article Assessment}

The top-cited audit shows that article access itself is already a limiting
condition for reproducibility assessment. Of the 100 citation-ranked
KNIME-matching records, 79 have a local PDF recorded in the compact audit
report. This means that the automated middle of the pipeline can evaluate most,
but not all, top-cited records from local article text and extracted
references. The remaining records stay in the denominator, because excluding
them would overstate the assessability of the literature.

The URL and reference-page stages add a more precise view of workflow
availability than article text alone. The compact report contains
workflow-relevant linked resources for 22 article records, with 68 linked
resources retained after filtering generic article pages, datasets, software
documentation, and malformed or unrelated URLs. This result is important
methodologically: the clearest evidence for whether a reader can obtain a
workflow often appears on the referenced page itself, not in the sentence that
cites the URL.

The LLM reference-classification step reduced the URL set to resources that
could plausibly support workflow retrieval. In the compact report, most retained
workflow links are classified as downloaded workflow artifacts, with only two
manual-check resources and a small number of landing-page or direct-workflow
references. The broader workflow inventory keeps 31 article records as
download candidates, because it preserves retrieval attempts even when the
compact article flag is later stricter. This separation is useful: it lets the
audit distinguish "the article or reference suggests a workflow may be
available" from "a workflow artifact was actually obtained locally."

The supplementary LLM flag step gives the publication-level reporting pattern
around those workflow links. Seventy-five records use KNIME as a workflow,
tool, interface, or implementation context. Most of these KNIME-use records
describe workflow steps or nodes in text, and many show workflow screenshots or
figures. However, much smaller subsets report a KNIME version, a downloadable
workflow, extension dependencies, or an extension installation source. This
indicates that visual or narrative workflow presentation is more common than
the executable and environment context needed for later reproduction.

The workflow-download stage converts reported or linked workflow evidence into
local artifact evidence. Eighteen article records yielded workflow files,
archives, or workflow directories. Four were obtained automatically and
fourteen required manual download. This confirms the semi-automatic nature of
the boundary step: scripted retrieval can find some artifacts, but protected
pages, stale links, repository layouts, and supplement portals still require
human inspection.

Table~\ref{tab:top-cited-assessment} summarizes the statistics-ready fields in
the compact article audit report. Data and code signals are recorded separately
from workflow-artifact evidence. We do not verify the availability of reported
data or code in this table; we record whether the article and linked evidence
report those resources.

\begingroup
\scriptsize
\renewcommand{\arraystretch}{1.18}
\begin{longtable}{|%
  >{\raggedright\arraybackslash}p{0.34\textwidth}|
  r|
  r|
  >{\raggedright\arraybackslash}p{0.37\textwidth}|}
\caption{Flag summary for the top-cited article audit. Shares use all
100 records in the registry as the denominator.}
\label{tab:top-cited-assessment}\\
\toprule
Audit category or flag & Count & Share of 100 & Interpretation \\
\midrule
\endfirsthead
\toprule
Audit category or flag & Count & Share of 100 & Interpretation \\
\midrule
\endhead
\bottomrule
\endfoot
Audit records & 100 & 100.0\% & Most-cited KNIME-matching article records selected from OpenAlex \\
Records with local PDF & 79 & 79.0\% & Source PDF recorded in the compact audit report \\
\midrule
Uses KNIME & 75 & 75.0\% & KNIME used as a workflow, tool, interface, or implementation context \\
Workflow or nodes described in text & 68 & 68.0\% & Named workflow steps, nodes, modules, or components recorded \\
Workflow screenshots or figures & 49 & 49.0\% & Workflow shown visually in article or supplement \\
Reports KNIME version & 18 & 18.0\% & Specific KNIME version value found \\
Reports downloadable workflows & 22 & 22.0\% & Executable workflow files provided or linked \\
Reports extension/plugin dependencies & 23 & 23.0\% & KNIME extension or node-library dependency reported \\
Reports extension installation source & 4 & 4.0\% & Update site, project URL, or installation source reported \\
\midrule
Provides code or scripts & 20 & 20.0\% & Code repository, scripts, or software code reported \\
Reports input-data availability & 66 & 66.0\% & Data availability stated, with or without direct URL \\
Direct input-data resource & 20 & 20.0\% & Direct data URL or resource pointer recorded \\
\end{longtable}
\endgroup

\begingroup
\scriptsize
\renewcommand{\arraystretch}{1.18}
\begin{longtable}{|%
  >{\raggedright\arraybackslash}p{0.62\textwidth}|
  r|
  r|}
\caption{Workflow-reporting and retrieval signals among the 75 assessed KNIME-use cases.}
\label{tab:knime-use-assessment}\\
\toprule
Audit flag & Count & Share of KNIME-use cases \\
\midrule
\endfirsthead
\toprule
Audit flag & Count & Share of KNIME-use cases \\
\midrule
\endhead
\bottomrule
\endfoot
Uses KNIME & 75 & 100.0\% \\
Workflow or nodes described in text & 68 & 90.7\% \\
Workflow screenshots or figures & 49 & 65.3\% \\
Reports KNIME version & 18 & 24.0\% \\
Reports extension/plugin dependencies & 23 & 30.7\% \\
Reports extension installation source & 4 & 5.3\% \\
Reports downloadable workflows & 22 & 29.3\% \\
Articles with successfully downloaded workflows & 18 & 24.0\% \\
\end{longtable}
\endgroup

For workflow retrieval, the authoritative data source is the workflow inventory
derived from the compact audit report and the local workflow directories.
Table~\ref{tab:workflow-retrieval-results} summarizes the recorded
artifact-retrieval outcomes used for manuscript inspection. In this table, the
\emph{Success} column marks whether a KNIME workflow artifact was actually
obtained: \textbf{X} means that a workflow file, archive, or workflow directory
was retrieved, while \textbf{-} means that no KNIME workflow artifact was
obtained or confirmed. Failed or incomplete retrievals are recorded separately
from article-level claims that a workflow was reported or linked.

Table~\ref{tab:workflow-opening-results} then reports the manual current-KNIME
opening and execution checks for the downloaded article records. Here, the
\emph{Executed} column has a stricter meaning: \textbf{X} means that at least
one workflow for the article opened and executed successfully in the tested
environment, while \textbf{-} means that the workflow opened but did not have a
confirmed successful end-to-end run, or failed during execution. Five article
records had at least one workflow execute successfully in the recorded manual
checks: PAINS, Webinar Pricing Analytics, ImageJ ecosystem integration,
high-content organelle trafficking, and GediNET
\cite{Saubern2011PAINS,VillarroelOrdenes2021KNIMEHub,Dietz2020ImageJKNIME,Pal2018OrganelleKNIME,Qumsiyeh2022GediNET}.
Other opened workflows did not have a confirmed successful end-to-end run in
this manual check. This does not show that those workflows are unreproducible:
they may require additional configuration, dependencies, or repair before a
successful run. Across the opened workflows, the screenshots also show
deprecated or legacy nodes, but we treat this as qualitative compatibility
evidence here rather than counting visible deprecated nodes. The screenshots
are archived under each workflow directory's \texttt{opened/} subdirectory in
\path{data/original/workflows/}.

\begingroup
\scriptsize
\renewcommand{\arraystretch}{1.18}
\begin{longtable}{|%
  >{\centering\arraybackslash}p{0.04\textwidth}|
  >{\raggedright\arraybackslash}p{0.23\textwidth}|
  >{\centering\arraybackslash}p{0.09\textwidth}|
  >{\raggedright\arraybackslash}p{0.54\textwidth}|}
\caption{Workflow retrieval results for article records with attempted or
recorded artifact retrieval outcomes.}
\label{tab:workflow-retrieval-results}\\
\toprule
\# & Article record & Success & Retrieval note \\
\midrule
\endfirsthead
\toprule
\# & Article record & Success & Retrieval note \\
\midrule
\endhead
\bottomrule
\endfoot
1 &
PAINS workflows, DOI \path{10.1002/minf.201100076} &
\textbf{X} &
Automated myExperiment page and download requests returned HTTP 403 HTML, but
we manually obtained both RDKit and Indigo workflow ZIP archives. Both
contain \texttt{workflow.knime} files. \\
2 &
Chemical data curation workflow, DOI \path{10.1186/s13321-018-0315-6} &
\textbf{X} &
GitHub master archive was downloaded after retry and passed ZIP integrity
checking. KNIME workflow files were found in the archive. \\
3 &
ASD genes prediction workflow, DOI \path{10.1371/journal.pone.0208626} &
\textbf{X} &
GitHub archive passed ZIP integrity checking. An inner KNIME workflow archive
was extracted to a workflow directory. \\
4 &
Bitterness/sweetness classifier workflow, DOI \path{10.3389/fchem.2018.00093} &
\textbf{-} &
The original project page redirected to VirtualTaste, and the redirected page
returned HTTP 404. No workflow archive was retrieved. \\
5 &
Medicinal chemistry filters workflow, DOI \path{10.3390/encyclopedia3020035} &
\textbf{X} &
GitLab archive passed ZIP integrity checking and contained KNIME workflow
files. One KNIME Hub URL returned a page, while an older generic Hub URL
returned 404 JSON. \\
6 &
MS-ready structures repository, DOI \path{10.1186/s13321-018-0299-2} &
\textbf{-} &
The GitHub archive passed ZIP integrity checking, but the extracted repository
contained only \texttt{LICENSE} and \texttt{README.md}. No \texttt{.knwf} or
\texttt{workflow.knime} file was found. \\
7 &
KNIME-OMICS reference, DOI \path{10.1016/j.jprot.2016.08.002} &
\textbf{-} &
The referenced GitHub organization page returned HTTP 404. No workflow archive
was retrieved. \\
8 &
QSAR tools workflow, DOI \path{10.1080/07391102.2018.1456975} &
\textbf{-} &
The referenced \path{teqip.jdvu.ac.in/QSAR_Tools/} URL failed DNS resolution.
No workflow archive was retrieved. \\
9 &
Webinar Pricing Analytics workflow, DOI \path{10.1016/j.jbusres.2021.08.036} &
\textbf{X} &
We manually downloaded the KNIME workflow file and added local KNIME
opening screenshots. \\
10 &
Verhaar scheme workflow, DOI \path{10.1016/j.chemosphere.2015.06.009} &
\textbf{-} &
The article reports a supplementary KNIME workflow. Standard Elsevier
supplement URLs resolved as Excel files, but no direct KNIME workflow archive
URL was confirmed. \\
11 &
Spontaneous tail coiling workflow, DOI \path{10.1016/j.ntt.2020.106918} &
\textbf{X} &
We manually downloaded the KNIME workflow file and added local KNIME
opening screenshots. \\
12 &
Caco-2 permeability prediction workflow, DOI \path{10.3390/pharmaceutics14101998} &
\textbf{X} &
We manually downloaded the KNIME workflow file and added local KNIME
opening screenshots. \\
13 &
KniMet workflow, DOI \path{10.1007/s11306-018-1349-5} &
\textbf{X} &
The Zenodo DOI resolved to record 1196407. \texttt{KniMet-master.zip} was
downloaded and contains \texttt{KniMet.knwf}. \\
14 &
NGS sample workflow, DOI \path{10.1093/bioinformatics/btr478} &
\textbf{X} &
We manually downloaded the myExperiment archive after automated requests
were blocked by HTTP 403. The ZIP contains a KNIME workflow directory with
saved data. \\
15 &
ImageJ ecosystem workflows, DOI \path{10.3389/fcomp.2020.00008} &
\textbf{X} &
Six KNIME Hub artifact metadata endpoints returned signed URLs and the
corresponding \texttt{.knwf} files were downloaded. One recorded reference
appears stale or unavailable. \\
16 &
Multi-template matching repository, DOI \path{10.1186/s12859-020-3363-7} &
\textbf{-} &
The GitHub archive was downloaded, but it contained documentation/images only
and no \texttt{.knwf}, \texttt{workflow.knime}, or KNIME workflow archive. A
guessed NodePit URL returned HTTP 404. \\
17 &
GediNET workflows, DOI \path{10.1038/s41598-022-24421-0} &
\textbf{X} &
GitHub master archive was downloaded and contains two \texttt{.knwf} files.
The recorded KNIME Hub short link could not be retrieved. \\
18 &
Nuclear receptor ligand workflow, DOI \path{10.1002/minf.201400188} &
\textbf{-} &
The article reports supplementary workflow material, but Wiley article and
guessed supplementary download endpoints returned Cloudflare challenge pages.
Crossref metadata did not expose a supplementary workflow URL. \\
19 &
High-content organelle trafficking workflow, DOI \path{10.1038/sdata.2018.241} &
\textbf{X} &
The figshare collection DOI resolved through the figshare API. The default
KNIME workflow file was downloaded as \texttt{KNIME workflow.zip}. \\
20 &
Knime4Bio NOTCH2 workflow, DOI \path{10.1093/bioinformatics/btr554} &
\textbf{X} &
The article reports that the workflow was posted on myExperiment. The local
manual artifact is \texttt{The\_NOTCH2\_workflow-v1.zip}, which contains a
KNIME workflow directory. \\
\end{longtable}
\endgroup

\begingroup
\scriptsize
\renewcommand{\arraystretch}{1.18}
\begin{longtable}{|%
  >{\centering\arraybackslash}p{0.04\textwidth}|
  >{\raggedright\arraybackslash}p{0.25\textwidth}|
  >{\centering\arraybackslash}p{0.09\textwidth}|
  >{\raggedright\arraybackslash}p{0.51\textwidth}|}
\caption{Manual opening and execution results for downloaded KNIME workflow
artifacts.}
\label{tab:workflow-opening-results}\\
\toprule
\# & Article record & Executed & Manual KNIME result \\
\midrule
\endfirsthead
\toprule
\# & Article record & Executed & Manual KNIME result \\
\midrule
\endhead
\bottomrule
\endfoot
1 &
PAINS workflows, DOI \path{10.1002/minf.201100076} &
\textbf{X} &
RDKit and Indigo workflow ZIPs were tested. At article-record level, at least
one workflow opened and executed successfully. The separate Indigo workflow
opened but was blocked by a missing Indigo KNIME Integration node. \\
2 &
Chemical data curation workflow, DOI \path{10.1186/s13321-018-0315-6} &
\textbf{-} &
The CompTox rebuilt workflow file opened, but execution failed inside a
metanode during the manual test. \\
3 &
ASD genes prediction workflow, DOI \path{10.1371/journal.pone.0208626} &
\textbf{-} &
The extracted ASD genes prediction workflow opened, but execution failed in an
R node because required R packages and functions were unavailable in the test
environment. \\
4 &
Medicinal chemistry filters workflow, DOI \path{10.3390/encyclopedia3020035} &
\textbf{-} &
The medicinal chemistry filters workflow file opened, but a successful
end-to-end execution was not established in the available notes. \\
5 &
Webinar Pricing Analytics workflow, DOI \path{10.1016/j.jbusres.2021.08.036} &
\textbf{X} &
The Webinar Pricing Analytics workflow file opened and executed successfully
in the current local KNIME environment. \\
6 &
Spontaneous tail coiling workflow, DOI \path{10.1016/j.ntt.2020.106918} &
\textbf{-} &
The spontaneous tail coiling measurement workflow file opened in the current
local KNIME environment, but a successful end-to-end execution was not
established in the available notes. \\
7 &
Caco-2 permeability prediction workflow, DOI \path{10.3390/pharmaceutics14101998} &
\textbf{-} &
The manual Caco-2 prediction \texttt{.knwf} file opened in the current local
KNIME environment, but a successful end-to-end execution was not established
in the available notes. \\
8 &
KniMet workflow, DOI \path{10.1007/s11306-018-1349-5} &
\textbf{-} &
The KniMet workflow from the Zenodo archive opened in the current local KNIME
environment, but successful execution was not established in the available
notes. \\
9 &
NGS sample workflow, DOI \path{10.1093/bioinformatics/btr478} &
\textbf{-} &
The NGS sample workflow archive with saved data opened in the current local
KNIME environment, but successful execution was not established in the
available notes. \\
10 &
ImageJ ecosystem workflows, DOI \path{10.3389/fcomp.2020.00008} &
\textbf{X} &
Six downloaded KNIME Hub \texttt{.knwf} files opened and executed normally in
our test environment after installing additional required KNIME/ImageJ-related
plugins. \\
11 &
GediNET workflows, DOI \path{10.1038/s41598-022-24421-0} &
\textbf{X} &
The GediNET GitHub archive opened in the current local KNIME environment and
the workflow was run successfully in the manual check. \\
12 &
High-content organelle trafficking workflow, DOI \path{10.1038/sdata.2018.241} &
\textbf{X} &
\texttt{KNIME workflow.zip} opened and executed normally in our test
environment after installing additional required plugins. \\
13 &
Knime4Bio NOTCH2 workflow, DOI \path{10.1093/bioinformatics/btr554} &
\textbf{-} &
\texttt{The\_NOTCH2\_workflow-v1.zip} opened in the current local KNIME
environment, but a successful end-to-end execution was not established in the
available notes. \\
\end{longtable}
\endgroup

\subsection{Repository-Level Analysis}

The repository-mining results explain why weak workflow reporting becomes more
costly over time. Deprecated nodes are not an isolated metadata corner case in
the KNIME source tree. At the earliest available source-code baseline,
2018-04-03, 193 of 1301 registered ordinary nodes were marked as deprecated, or
14.83\% (Table~\ref{tab:selected-snapshots}). By the KNIME Analytics Platform
5.0.0 source-date anchor on 2023-02-22, the count had increased to 433 of 1442
ordinary nodes, or 30.03\%. The current local source-date snapshot on June 28,
2026, contains 502 deprecated ordinary nodes out of 1506, or 33.33\%. These
figures show that the deprecated-node share approximately doubled between the
2018 baseline and the 2026 local snapshot.

The annual checkpoint table gives the same trend at regular intervals
(Table~\ref{tab:annual-snapshots}). The number of repositories with checkout
commits rises from 44 in 2018 to 90 in 2026, so early and late totals should
not be compared as if the mined source corpus were fixed. Nevertheless, the
deprecated-node share rises from 14.83\% in 2018 to 33.47\% at the 2026 annual
checkpoint. Hidden ordinary nodes remain much less frequent than deprecated
ordinary nodes, but they increase from 2 in 2018 to 24 in 2026. Deprecated
dynamic nodesets appear from 2020 onward and remain at 2 in the annual
checkpoints through 2026.

\begin{table}
\centering
\caption{Selected KNIME repository-mining snapshots.}
\label{tab:selected-snapshots}
\scriptsize
\begin{tabular}{lp{0.34\textwidth}rrr}
\toprule
Date & Version context & Nodes & Deprecated nodes &
\shortstack{Deprecated\\share} \\
\midrule
2018-04-03 & KNIME Analytics Platform 3.5.3 source-date anchor & 1301 &
193 & 14.83\% \\
2019-12-05 & KNIME Analytics Platform 4.1.0 source-date anchor & 1191 &
227 & 19.06\% \\
2023-02-22 & KNIME Analytics Platform 5.0.0 source-date anchor & 1442 &
433 & 30.03\% \\
2026-03-03 & KNIME Analytics Platform 5.11.0 source-date anchor & 1503 &
503 & 33.47\% \\
2026-06-28 & Current local source-date snapshot & 1506 & 502 & 33.33\% \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}
\centering
\caption{Annual KNIME repository-mining checkpoint statistics.}
\label{tab:annual-snapshots}
\scriptsize
\begin{tabular}{llrrrrrr}
\toprule
Year & Snapshot date & Repositories & Nodes &
\shortstack{Deprecated\\nodes} &
\shortstack{Deprecated\\nodesets} &
\shortstack{Hidden\\nodes} &
\shortstack{Deprecated\\share} \\
\midrule
2018 & 2018-04-03 & 44 & 1301 & 193 & 0 & 2 & 14.83\% \\
2019 & 2019-01-01 & 45 & 1500 & 248 & 0 & 2 & 16.53\% \\
2020 & 2020-01-01 & 54 & 1200 & 228 & 2 & 2 & 19.00\% \\
2021 & 2021-01-01 & 60 & 1366 & 398 & 2 & 5 & 29.14\% \\
2022 & 2022-01-01 & 67 & 1424 & 386 & 2 & 8 & 27.11\% \\
2023 & 2023-01-01 & 72 & 1428 & 433 & 2 & 10 & 30.32\% \\
2024 & 2024-01-01 & 80 & 1434 & 436 & 2 & 9 & 30.40\% \\
2025 & 2025-01-01 & 86 & 1488 & 446 & 2 & 23 & 29.97\% \\
2026 & 2026-01-01 & 90 & 1509 & 505 & 2 & 24 & 33.47\% \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}
\centering
\caption{Annual KNIME repository-mining transition statistics.}
\label{tab:annual-transitions}
\scriptsize
\begin{tabular}{lrrrrr}
\toprule
Year & Added & Removed &
\shortstack{Newly\\deprecated} &
\shortstack{No longer\\deprecated} &
\shortstack{Category\\changed} \\
\midrule
2018 & -- & -- & -- & -- & -- \\
2019 & 194 & 22 & 16 & 0 & 6 \\
2020 & 11 & 0 & 0 & 0 & 4 \\
2021 & 140 & 1 & 80 & 0 & 111 \\
2022 & 69 & 6 & 17 & 29 & 17 \\
2023 & 56 & 4 & 61 & 0 & 67 \\
2024 & 17 & 7 & 4 & 0 & 3 \\
2025 & 36 & 0 & 9 & 0 & 1 \\
2026 & 52 & 2 & 60 & 0 & 3 \\
\bottomrule
\end{tabular}
\end{table}

The transition table indicates that node deprecation changes occur in bursts
rather than at a constant rate (Table~\ref{tab:annual-transitions}). The largest
annual increases in newly deprecated node identities occur at the 2021, 2023,
and 2026 checkpoints, with 80, 61, and 60 newly deprecated node keys,
respectively. The 2022 checkpoint is the only annual checkpoint in which a
substantial number of common node keys are no longer marked deprecated: 29.
Category-path changes are also concentrated in a few years, especially 2021
and 2023. Because these transition counts use metadata-level node identity keys,
they should be read as evidence of source metadata evolution, not as direct
evidence that published workflows fail to open or execute.

These repository-level results directly address the second research question:
how KNIME node deprecation evolved over time. They also provide necessary
context for the workflow-level questions, because a published workflow that
depends on older node identities may encounter deprecated, hidden, removed, or
migrated components in a later KNIME version. However, the tables do not answer
how often uploaded or executed workflows use deprecated nodes. That question
requires workflow-level corpora and execution traces beyond repository
metadata.

Taken together, the evidence gives the intended picture of the study. First,
KNIME is visible in a broad and sustained scientific literature. Second,
influential papers that use or mention KNIME rarely provide the full
workflow-preservation package needed for later reproduction: executable
workflow files, KNIME versions, extension context, data, and code. Third, the
workflow experiment shows the practical consequence of this gap: 18 article
records yielded workflow artifacts or workflow directories, workflows from all
18 opened in the current local KNIME environment, and five article records had
at least one workflow execute successfully. Fourth, KNIME's source metadata
continues to evolve, and roughly one third of ordinary registered nodes are
marked deprecated in recent source snapshots. The result is not a claim that
all KNIME workflows fail. The result is a reproducibility risk: as KNIME
evolves, papers that preserve only workflow images or textual descriptions
become harder to diagnose, migrate, and reproduce.

\section{Discussion}

To our knowledge, this is the first focused empirical study connecting KNIME
workflow reporting in scientific publications with KNIME source-code evolution.
The contribution is intentionally narrower than a full reproducibility
benchmark. We do not estimate the failure rate of all published KNIME
workflows. Instead, we show a reproducibility risk pattern: KNIME is used in a
visible scientific literature, influential papers often do not publish
executable workflow artifacts, successful current-version execution was
confirmed for five article records in the recorded manual checks, and the
KNIME node ecosystem has accumulated a substantial deprecated-node surface.

The results suggest three practical implications. First, a published workflow
image is not an executable artifact. Several highly cited papers provide
figures or textual descriptions of connected KNIME nodes, but this is not
enough to recover node settings, extension versions, input paths, or migration
requirements. Second, reporting a KNIME version is useful but incomplete.
Workflows may also depend on third-party extension update sites and node
libraries, as shown by the PAINS case where the RDKit workflow could be
updated and executed while the Indigo workflow was blocked by a missing
extension. Third, repository-level deprecation is a useful warning signal. It
does not prove that a workflow fails, but it identifies a compatibility surface
that grows as the platform evolves. Without careful workflow preservation and
version reporting, this growth makes old workflows harder to inspect and
repair.

The next development of this study should move from paper-level evidence to
workflow-level evidence. One direction is to create a larger corpus of public
KNIME workflows from supplements, myExperiment, KNIME Hub, GitHub, and other
repositories, then test whether they open in current KNIME versions and which
deprecated, hidden, migrated, or missing nodes they contain. A second direction
is to link deprecated nodes to migration metadata and replacement nodes in the
KNIME source tree, so that workflow warnings can distinguish low-risk
deprecations from nodes that require unavailable extensions or manual repair.
A third direction is to preserve reproducibility environments for selected
case studies by recording operating system, KNIME version, installed
extensions, update sites, input data, import warnings, execution errors, and
output differences.

Tools and services outside the present evidence base can provide the next
layer of usage evidence. \texttt{knime2py} can help identify KNIME workflows that users
try to translate or operationalize outside KNIME \cite{KaplanKnime2py2026},
and \texttt{k2pweb.org} can provide anonymized aggregate evidence about real uploaded
or executed workflows \cite{K2PWeb2026}. These
sources would address a question that the current paper deliberately leaves
open: how often researchers and analysts encounter deprecated KNIME nodes in
practice, not only how often deprecated nodes appear in the source repository
or in selected published artifacts.

\bibliographystyle{splncs04}
\bibliography{references}

\end{document}
